---
title: "Detecting the 2022 Hunga Tonga eruption pressure wave in PWS data"
author: "Neil Saunders"
date: "compiled `r Sys.time()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(jsonlite)
library(lubridate)

theme_set(theme_bw())
```

# Summary
The Hunga Tonga eruption of January 15 2022 sent a pressure wave around the earth.

This document looks for the wave in atmospheric pressure data from local personal weather stations.

# Data
The function `json2df` takes a PWS ID, date and API key, retrieves data for that date in JSON format and returns a data frame.

```{r json2df}
json2df <- function(id, date, key) {
  u <- paste0("https://api.weather.com/v2/pws/history/all?stationId=", id, "&format=json&units=m&date=", date, "&apiKey=", key)
  u %>% 
    fromJSON() %>% 
    .$observations %>% 
    as_tibble() %>% 
    unnest(cols = c(metric))
}
```


```{r get-data}
apikey <- Sys.getenv('WUNDERGROUNDID')

stations <- c("ISYDNE1319", "ISYDNE290", "INSWMOUN16", "ISYDNEY156", 'ISYDNE1793', "ISYDNE1727")

wudata <- lapply(stations, function(x) json2df(x, "20220115", apikey)) %>% 
  bind_rows()
```

# Visualisation
The pressure wave is visible as a spike beginning just after 6 PM.

The eruption occurred at around 3:15 PM Sydney time. Sydney to Tonga is approximately 3 500 km. Assuming the standard speed of sound (343 m s$^-1$), we expect a travel time of `r 3500 / 0.343 / 60 / 60` hours. This is consistent with the time of the spike.

```{r plot-data}
wudata %>% 
  mutate(dt = ymd_hms(obsTimeLocal)) %>%
  ggplot(aes(dt, pressureMax)) + 
  geom_line() + 
  facet_grid(stationID ~ ., scales = "free_y") + 
  scale_x_datetime(breaks = "3 hours", date_labels = "%H:%M") +
  labs(x = "Time",
       y = "Max Pressure (hPa)",
       title = "Air pressure at 6 personal weather stations in Hornsby Heights",
       subtitle = "January 15 2022") +
  theme(strip.text = element_text(size = 6))
```

