---
title: "Winter rainfall in Sydney 2017"
author: "Neil Saunders"
date: "compiled `r Sys.time()`"
output: 
  html_document: 
    keep_md: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = FALSE, message = FALSE, warning = FALSE, fig.path = "../../figures/")
library(tidyverse)
library(lubridate)

theme_set(theme_bw())

# HH data from the rwundergound package
wu_hh <- readRDS("../../data/wu_hh.rds")
```

# Introduction
It feels like the winter and spring of 2017 in Sydney were exceptionally dry. Does data back up this feeling?

# Read and transform the data
Unfortunately, `rwunderground::history_range()` is returning NA values for many variables, including rainfall:

```{r hh-data}
wu_hh %>%
  glimpse()
```

So data for the Hornsby personal weather station (INEWSOUT879) was copy-pasted from the Weather Underground website into a text file, so requires some additional processing.

```{r read-data}
hornsby <- read_delim("../../data/hornsby.txt", "\t", escape_double = FALSE, 
                      col_names = FALSE, trim_ws = TRUE)

hornsby <- hornsby %>%
  mutate(date = c(paste("2017", "01", 1:31, sep = "-"), paste("2017", "02", 1:28, sep = "-"), paste("2017", "03", 1:31, sep = "-"), paste("2017", "04", 1:30, sep = "-"), paste("2017", "05", 1:31, sep = "-"), paste("2017", "06", 1:30, sep = "-"), paste("2017", "07", 1:31, sep = "-"), paste("2017", "08", 1:31, sep = "-"), paste("2017", "09", 1:16, sep = "-"), paste("2017", "09", 21:30, sep = "-"), paste("2017", "01", 1:9, sep = "-")),
         date = as.Date(date)) %>%
  rename(day = X1, temp_high = X2, temp_mean = X3, temp_low = X4, dew_high = X5, 
         dew_mean = X6, dew_low = X7, humidity_high = X8, humidity_mean = X9, humidity_low = X10,
         windspeed_high = X11, windspeed_mean = X12, windspeed_gust = X13, pressure_high = X14,
         pressure_mean = X15, pressure_low = X16, precip_total = X17)
```

# Plot the data
```{r plot-data}
hornsby %>% 
  ggplot(aes(date, precip_total)) + 
    geom_line(color = "#0089a8") + 
    labs(x = "Date", y = "precipitation (mm)", 
         title = "Precipitation Jan - Oct 2017", 
         subtitle = "source: WU INEWSOUT879") + 
    theme_light() + 
    scale_x_date(date_breaks = "1 month", date_labels = "%b")
```

# Conclusion
Very little significant rainfall was measured between April and the beginning of October. So yes: our feeling that "it hasn't rained for months" is correct.